### 2019-06-03 Declaration/declaration.tex
Updated: Changed info to Newcastle University and me as author

% ******************************* Thesis Declaration ***************************

\begin{declaration}

I hereby declare that except where specific reference is made to the work of 
others, the contents of this dissertation are original and have not been 
submitted in whole or in part for consideration for any other degree or 
qualification in this, or any other university. This dissertation is my own 
work and contains nothing which is the outcome of work done in collaboration 
with others, except as specified in the text and Acknowledgements. This 
dissertation contains fewer than 80,000 words including appendices, 
bibliography, footnotes, tables and equations.

% Author and date will be inserted automatically from thesis.tex \author \degreedate

\end{declaration}