### 2020-01-10 ChapterConclusion/chapterConclusion.tex
Updated: Expanded Futher Work

\chapter{Conclusions and Future Work}
An approach to networked real-time physics simulations that is scalable and alleviates the processing limitation of a single server has been presented. Only open-source software has been used in our approach and our algorithm has been developed in a way that is agnostic to any specific application technology.

Our experiments establish that the approach is scalable, as demonstrated by the addition of servers improving the performance of the system when simulating an increasingly large number of objects. This study has demonstrated that a standard real-time physics engine (in this case, PhysX) may be incorporated into our scalable real-time physics system and achieve performance that is acceptable for real-time distributed simulations such as networked games.

We have measured and evaluated the correctness of collisions in AP. Our experiments establish that collisions remain correct in AP, even when objects interact while intersecting the boundary. This has been shown to be true while increasing each of the following factors: speed; latency and frame-time. Collisions remain correct up to the tolerance values for each factor and begin to show errors as soon as the tolerance is exceeded, demonstrating the correctness of the aura calculation used in AP. In addition the effects of packet loss on the system have also been explored and the limits of AP with increasing packet loss have been shown. This paper demonstrates that collisions of rigid-bodies can be handled correctly in a scalable, distributed real-time physics system, even when those collisions take place on or near a server-region boundary.

%Thrashing, if we do it?

Our continuing work will be to carry out experiments to determine the effects that changing each of the aura tolerances has on the performance and scalability of AP. Future work will enable each tolerance value to be dynamic and the aura size could adapt with the changing conditions of the system. This would improve performance when factors are low and maintain correctness when factors are high.

Future experiments may also involve the decoupling of ``update rate'' from ``tick rate'' as described in Section \ref{update-rate-test-values}. As some video games opt to maintain a high internal tick rate to improve simulation accuracy but use a lower or variable update rate to save bandwidth when communicating with networked peers, it would be of interest to examine how doing so would affect the correctness of aura projection in these configurations.
Additionally, on-demand dynamic load balancing will be explored to enable improvements in scalability. This will leverage run-time resource (re)allocation in order to maintain efficient distribution of cloud-based processing facilities in response to changes in the simulated environment. Simulations of objects that can freely interact with one another with a high degree of fidelity are never uniform, and therefore on-line statistical modelling and analysis will be fundamental to predicting where additional computational units are required. This will necessitate a transition from fixed world region boundaries to dynamic re-partitioning, which we expect will improve performance by reducing the number of auras being projected and the number of objects being migrated, leading to an overall reduction in bandwidth usage.

We will also be researching the use of modern cloud technologies such as containers and container orchestration in order to facilitate efficient provisioning and management of additional compute resources within a distributed physics simulation. The use of lightweight containers versus virtual machine instances may allow for more precise control of compute resource allocation and faster on-demand provisioning, which would lend itself well to streamed gaming applications.

In order for the simulation to be truly scalable, new regions will need to be created during run-time (either dividing existing regions or expanding the simulation area). New nodes will need to be launched during run-time and a method for re-dividing up the workload between the new and existing nodes will be required.

AP works on the principle that any two bodies that may be interacting are always being simulated on the same server. An alternative technique to this would be Cross-Boundary Interaction, as discussed in \ref{ConsideredSolutions}, allowing for bodies to interact across region boundaries i.e. objects that are being simulated on two different servers interacting over the network. There are likely various advantages and disadvantages of Cross-Boundary Interaction compared to Aura Projection, which could be explored and evaluated. The two solutions may have better performance in different situations (such as a large cluster on either side of a boundary) and a solutions which uses both techniques in different circumstances could be developed.

An important feature of a physics engine is the ability to query the state of the simulation. A typical example of this would be a ray-cast, in which a ray is cast between two points in the simulation space to find intersecting bodies. %Ray-casts are commonly used in first-person shooting games, i.e. a ray is cast from the player's weapon to a given distance, and objects that intersect the ray are returned from the query.
Distributed physics adds an extra level of complexity to this as the query may concern spaces and/or objects being simulated on different servers. In the case of spatially-partitioned solutions, the query could overlap the region boundary and in order for the query to be completed, a round-trip of messages between the two servers would be required. If the query is blocking/synchronous (execution cannot continue until the query is complete), this would introduce a large amount of wait time (at least double the network latency). Another option would be to use asynchronous queries, the results of which get returned using a function callback as soon as the results have been received. However this is more complex than a query on a centralised simulation which can return the results synchronously without any network delay.