### 2020-03-24 Acknowledgement/acknowledgement.tex
Updated: Acknowledgement and Dedication

% ************************** Thesis Acknowledgements **************************

\begin{acknowledgements}

I would like to thank my supervisor Dr. Graham Morgan for his continuous support and guidance throughout my PhD study and research.
I would like to acknowledge Dale Whinham for his improvements to the build and deploy system used in this project and for his conversion of some of the diagrams from the office whiteboard into InkScape. I would also like to acknowledge Carlos Guerrero Rodriguez for his contributions under my guidance to the visualiser, including the client-created replicas and static objects, client-defined server regions and reset features.

My sincere thanks to my parents and partner for their excellent proof-reading skills and support.

\end{acknowledgements}